---
title: "Biblical"
author: "Sharon Howard"
date: "2025-02-26"
execute: 
  message: false
  warning: false
  echo: false
---



```{r}
library(readxl)
#library(writexl)

#library(readtext)
library(janitor)
library(lubridate)
library(scales)

library(snakecase)
library(glue)
# library(knitr)
# library(kableExtra)

# library(XML)
# library(xml2)
# library(rvest)
# library(xslt)


library(ggmosaic)
library(ggthemes)

library(tidytext)
library(tidyverse)


# gg theme replacing minimal
theme_set(theme_gray())
theme_update(panel.background = element_rect(fill="#fafafa"), 
             strip.background = element_rect(fill="#f0f0f0"),
             axis.ticks = element_line(colour = 'grey'))



# do i want this? maybe. 
bible_list <- 
  tribble(~"vol", ~"name",
          "KJV", "King James",
          "GB", "Geneva Bible",
          "BB", "Bishops Bible",
          "BCP", "Book of Common Prayer"
          )

books_table_all_rds <-
  readRDS("~/r_projects/atb/outputs/bibl_books_table_all_20250226.Rds")


## unnest individual refs
books_table_all_refs <-
books_table_all_rds |>
   unnest_regex(ref, source, pattern = " +", to_lower = FALSE) |>
   relocate(ref) |>
   # add a rowid for each reference. id column is for <quote>
   rowid_to_column("ref_id") |>
   separate(ref, into = c("r1", "r2"), ",?_", remove = F, extra = "merge")



bcp_other_references <-
books_table_all_refs |>
  filter(str_detect(ref, "BCP,")) |>
  mutate(section = case_when(
    str_detect(r2, "Holy_Communion") ~ "Holy Communion",
    str_detect(r2, "^Burial_of_the_Dead") ~ "Burial",
    str_detect(r2, "Morning_Prayer") ~ "Morning Prayer",
    str_detect(r2, "Evening_Prayer") ~ "Evening Prayer",
    str_detect(r2, "Baptism") ~ "Baptism",
    r2 %in% c("The_Litany", "Prayer_for_Persons_Troubled_in_Mind", "Thanksgiving_for_Deliverance_from_Plague", "Thanksgiving_for_the_Restoration") ~ r2,
    str_detect(r2, "Prayers_at_Sea|Storms_at_Sea") ~ "At Sea",
    str_detect(r2, "^The_Collect") ~ "The Collect",
    str_detect(r2, "_Sick") ~ "The Sick",
    str_detect(r2, "Thanksgiving") ~ "Other Thanksgiving", 
    str_detect(r2, "Prayer") ~ "Other Prayer", 
    .default = "Other"
  )) |>
  mutate(section = str_replace_all(section, "_", " "))  |>
  mutate(vol = "BCP") |>
  mutate(book = "Liturgy") |>
  mutate(bibl = "BCP") |>
  select(ref_id, vol, book, bibl, section, src, type, text, id, ref, r2)


biblical_references <-
books_table_all_refs |>
  filter(!str_detect(ref, "BCP,") & !str_detect(ref, "Apocrypha")) |>
  separate(r2, into = c("book", "r3"), sep="_", extra = "merge") |>
  mutate(r3 = str_remove(r3, "^Psalm")) |>
  # keep distinct for the moment. ok i think refs look ok.
  mutate(book = case_when(
    book=="Psalter" ~ "Psalms",
    .default = book
  )) |>
  # fail to convert 
  # there are some ranges in c (only 4) plus "Susanna" which is book apocrypha - just lose that one
  separate(r3, into=c("chap", "v"), sep=":", fill="right") |>
  mutate(chap = str_remove(chap, "-.*$")) |>
  mutate(chap = parse_number(chap)) |>
  rename(vol=r1) |>
  mutate(vol = str_remove(vol, "#")) |>
# you need a label for NA vol... NS (not specified) ?
  mutate(vol = case_when(
    vol== "" ~ "NS",
    .default = vol
    )) |>
  mutate(bibl = case_when(
    vol=="BCP" ~ "BCP",
    .default = "Biblical"
  )) |>
  select(ref_id, vol, book, bibl, chap, v, src, type, text, id, ref)


bibl_bcp_references <-
bind_rows(
  biblical_references,
  bcp_other_references
) |>
  arrange(ref_id) 




#very plain text...
#mod spelling etc but no editorial additions (<supplied>) and also dropped <surplus>.
# no titles, no indexes, etc.

bookrem_txt_lines <- read_lines("~/my_docs/work/thornton/texts/transformations/txt/book_of_remembrances_mod_simplified_txt_just_txt.txt")
book1_txt_lines <- read_lines("~/my_docs/work/thornton/texts/transformations/txt/book_one_mod_simplified_txt_just_txt.txt")
book2_txt_lines <- read_lines("~/my_docs/work/thornton/texts/transformations/txt/book_two_mod_simplified_txt_just_txt.txt")
book3_txt_lines <- read_lines("~/my_docs/work/thornton/texts/transformations/txt/book_three_mod_simplified_txt_just_txt.txt")


bookrem_txt_words <-
bookrem_txt_lines |>
  enframe(name = NULL, value = "line") |>
  filter(line!="") |>
  unnest_tokens(word, line)

book1_txt_words <-
book1_txt_lines |>
  enframe(name = NULL, value = "line") |>
  filter(line!="") |>
  unnest_tokens(word, line)

book2_txt_words <-
book2_txt_lines |>
  enframe(name = NULL, value = "line") |>
  filter(line!="") |>
  unnest_tokens(word, line)

book3_txt_words <-
book3_txt_lines |>
  enframe(name = NULL, value = "line") |>
  filter(line!="") |>
  unnest_tokens(word, line)



books_txt_words <-
bind_rows(
  bookrem_txt_words |> mutate(src="Book Rem"),
  book1_txt_words |> mutate(src="Book 1"),
  book2_txt_words |> mutate(src="Book 2"),
  book3_txt_words |> mutate(src="Book 3")
)
```


It seemed to me that Book Rem might be over-represented for biblical citations (relative to its size), so I got word counts and biblical references (individual citations) and this shows each Book's percentage of the total for both measures. Book Rem *is* punching slightly above its weight, but Book 1 more so, and what's perhaps even more noticeable is how Book 3 is *under* represented. (Book 2 is about even.)


```{r}
bibl_bcp_references |>
  count(src, name="n_refs") |>
  inner_join(
    books_txt_words |>
  count(src, name="n_words"), by="src"
  ) |>
  # mutate(p = n_words/n_refs) |>
  mutate(tw = sum(n_words), tr= sum(n_refs)) |>
  # book word count as % of total words
  mutate(prop_of_total_words = n_words/tw*100) |>
  # book reference count as % of total references
  mutate(prop_of_total_references = n_refs/tr*100) |>
  pivot_longer(c(prop_of_total_words, prop_of_total_references), names_to = "p", values_to = "percent") |>
  ggplot(aes(p, percent, fill=src)) +
  geom_col() +
  #geom_text(aes(label=percent)) +
  scale_fill_ptol() +
  labs(y="%", x=NULL, fill="Book", title = "Compare Books' use of biblical references")

```

It's quite difficult to come up with effective ways to compare references because so many (bible) books are mentioned in the Books. This only includes the top 30. I've put BCP Psalms in with the rest of Psalms and bundled all the rest of BCP as "Liturgy".

```{r}
bibl_bcp_references |>
  # filter(book %in% c("Psalms", "Psalter")) |>
  # count(chap, v, book)
  #add_count(book, name="n_book") |>
  count(book, sort = T) |>
  # group_by(src) |>
  top_n(30, n) |>
  # ungroup() |>
  mutate(book=fct_reorder(book, n)) |>
  #mutate(book=reorder_within(short_label, pm, book_r))
  ggplot(aes(n, book)) +
  geom_col() +
  labs(title="Most frequently referenced books in the Books", x="references", y=NULL)
```

Comparison per book

```{r}
bibl_bcp_references |>
  # filter(book %in% c("Psalms", "Psalter")) |>
  # count(chap, v, book)
  add_count(book, name="n_book") |>
  count(book, src, n_book, sort = T) |>
  mutate(p = n_book/sum(n_book)*100) |>
  filter(p > 0.25) |>
  # group_by(src) |>
  # top_n(20, n) |>
  # ungroup() |>
  mutate(book=fct_reorder(book, n_book)) |>
  #mutate(book=reorder_within(short_label, pm, book_r))
  ggplot(aes(n, book)) +
  geom_col() +
  facet_wrap(~src, nrow=1, scales="free_x") +
  labs(title="Most frequently referenced books, by Book", x="references per book", y=NULL, caption = "books ordered by overall counts")
```

Because Psalms are so predominant, they make it difficult to see differences among the rest. So this is the same but with Psalms removed.

```{r}
bibl_bcp_references |>
  filter(!book %in% c("Psalms")) |>
  # count(chap, v, book)
  add_count(book, name="n_book") |>
  count(book, src, n_book, sort = T) |>
  mutate(p = n_book/sum(n_book)*100) |>
  filter(p > 0.25) |>
  # group_by(src) |>
  # top_n(20, n) |>
  # ungroup() |>
  mutate(book=fct_reorder(book, n_book)) |>
  #mutate(book=reorder_within(short_label, pm, book_r))
  ggplot(aes(n, book)) +
  geom_col() +
  facet_wrap(~src, nrow=1, scales="free_x") +
  labs(y=NULL, x="references per book", caption = "books ordered by overall counts", title="Most frequently referenced books, by Book, excluding Psalms")
```

Comparison of Biblical v BCP.

```{r}
bibl_bcp_references |>
  #filter(book !="Psalms") |>
  count(bibl, src) |>
  mutate(src=fct_rev(src)) |>
  ggplot(aes(n, fill=bibl, y=src)) +
  geom_col(position = "fill") +
  scale_fill_tableau() +
  scale_x_continuous(labels = percent_format()) +
  labs(y=NULL, fill="type", x="% of references", title = "Comparison of Biblical and BCP references per Book")
```


Comparison of reference types. 

```{r}
bibl_bcp_references |>
  count(type, src) |>
  filter(type !="biblical") |>
  mutate(src=fct_rev(src)) |>
  ggplot(aes(n, fill=type, y=src)) +
  geom_col(position = "fill") +
  scale_fill_tableau()+
  scale_x_continuous(labels = percent_format()) +
  labs(y=NULL, x="% of references", title = "Comparison of reference types per Book")
```





