---
title: "Mapping Alice Thornton's Books: Events"
author: "Sharon Howard"
execute: 
  message: false
  warning: false
  echo: false
toc: false
---

## About the Map

This is an interactive map made by [Sharon Howard](https://sharonhoward.org) for the [Alice Thornton's Books project](https://thornton.kdl.kcl.ac.uk/). 

The map shows places associated with narratives of [Events](https://thornton.kdl.kcl.ac.uk/posts/blog/2022-11-28-tei-events-thornton-books/) tagged by the project, with links to the project's [digital scholarly edition (DSE)](https://thornton.kdl.kcl.ac.uk/edition/?p0.lo=p.1&p0.vi=modern)




```{r}

get_shared_data <- function(){
  "~/r_projects/atb/shared_data/"  
}

library(readxl)
library(readtext)
library(janitor)
library(scales)

library(snakecase)
library(glue)

library(xml2)
library(rvest)

library(reactable)
library(htmltools)

library(tidytext)
library(tidyverse)

```


```{r}
#omitted atb_homes... need to drop refs to category


# this is a copy of the final xlsx in unlinked edinburgh onedrive folder at 26/2/25 
# is it safe to use the address column, or should i reconstruct from segments? i can't quite remember...

atb_places_wikidata_xlsx <-
  read_excel("~/r_projects/atb/shared_data/atb_places_v20230710.xlsx") |>
## add nls coords for the ones where wikidata isn't quite right...
  mutate(lat = case_when(
    !is.na(nls_lat) ~ nls_lat,
    .default = wd_lat
  )) |>
  mutate(long = case_when(
    !is.na(nls_long) ~ nls_long,
    .default = wd_long
  ))


atb_places <-
atb_places_wikidata_xlsx |>
  filter(level!="country") |> # keep counties (and the two sub-counties for now)
  mutate(lat=as.numeric(lat), long=as.numeric(long)) |>
  # make better address label. 
  mutate(label = case_when(
  # depends on empties being NA!
  # added level=non-country to deal with things like Irish Sea that are outside the usual hierarchy as they're not in a country
           level=="non-country" ~ placeName, 
           level=="country" ~ country,
           !is.na(placeName)  & is.na(settlement) ~ glue("{placeName}, {region}"),
           !is.na(placeName) ~ glue("{placeName}, {settlement}, {region}"),
           !is.na(settlement) ~ glue("{settlement}, {region}"),
           !is.na(region) ~ glue("{region}, {country}"),
           .default = placeName # in case anything is not caught above this should at least ensure blank or somethig.
         )) |>
  #left_join(atb_homes |> select(place_id, category, about), by="place_id") |>
  #mutate(category = if_else(!is.na(category), category, "other")) |>
  mutate(atb_places_url = glue("https://thornton.kdl.kcl.ac.uk/entities/?hi=place:{place_id}")) |>
  mutate(atb_places_link_html = glue("Search edition for mentions of <a href='{atb_places_url}' target='_blank'>{label}</a>")) 






# with tagged places only.

# # get the file names with full path. 
atb_evs_places_xml_files_folder <-
  list.files("~/r_projects/atb/shared_data/xml/evdivs_places/", pattern = "*.xml$", full.names = TRUE)

# but now you have to work out how to get the bits of xml you want... quicker right now to just read_lines? you know what code you need instead of having to look stuff up.

atb_fetch_evs_places_lines <-
  map(atb_evs_places_xml_files_folder, read_lines)


# need to check this still works and then work out to change the entities tags...
# seems good but you have a problem don't you. you can change all </?rs easily. and you can pick <rs type="whatever"
# but you can't diff </rs> for ppl names from </rs> for place names. feck. 
# you can't even assume that opening <rs will be on the same line as closing </rs> actually. 

# maybe i'll do it properly at some point lol
atb_evs_places_xml_txt <-
map(atb_fetch_evs_places_lines, enframe) |>
  bind_rows() |>
  mutate(b = cumsum(str_detect(value, "<TEI"))) |>
  filter(b != 0) |>
  group_by(b) |>
  mutate(t = cumsum(str_detect(value, "</?text>"))) |>
  mutate(xmlid = if_else(str_detect(value, "<TEI"), value, NA)) |>
  mutate(xmlid = str_match(xmlid, 'xml:id="([^"]+)"')[,2] ) |>
  fill(xmlid) |>
  filter(t==1 & !str_detect(value, "</?text|</?body")) |>
  mutate(value = str_trim(value)) |>
  filter(value !="") |>
  select(-t) |>
  mutate(d = cumsum(str_detect(value, "<div"))) |>
  mutate(id = if_else(str_detect(value, "<div"), value, NA)) |>
  mutate(id = str_match(id, 'xml:id="([^"]+)"')[,2]) |>
  ungroup() |>
  # make div atts more html-y, and then you can use css.
  mutate(value = str_replace(value, 'type="event"', 'class="atb-ev-mention"')) |>
  mutate(value = str_replace(value, 'xml:id="([^"]+)"', 'id="\\1"')) |>
  group_by(b, d) |>
  fill(id) |>
  ungroup()  |>
  # turn <rs into <em... that doesn't quite work.
  mutate(value = str_replace_all(value, "<rs", "<span class='hi_place'"), value = str_replace_all(value, "</rs", "</span")) |>
  group_by(xmlid, id) |>
  summarise(ev_text_html = glue_collapse(value, sep = " "), .groups = "drop_last") |>
  ungroup() |>
  separate(id, into = c("div_id", "ev"), sep = "\\.")





## events short titles have been updated i think but still a bit WIP
## are these and in events xlsx in correct date order? some look wrong in here. use the final event sort

# dates ugh. i see i have a lot of when/notbefore/from tei stuff going on
# just simplify to first year for now? use that for the filter, display can be more complex.

events_short_titles_xlsx <-
  read_excel(paste0(get_shared_data(), "atb_events_short_titles_20250606_ed.xlsx"), sheet = "events")

# don't think this includes parent events. 
events_xlsx <-
  read_excel(paste0(get_shared_data(), "atb_events_groups_ev_ids_20250408.xlsx"), sheet = "events")

# based on events standoff in atb, adding year.
events_dates <- 
events_xlsx |> 
  select(event_id, event_date) |>
  separate(event_date, into=c("start", "end"), sep="/|\\|", remove = F, fill = "right") |>
  mutate(date_type = case_when(
    is.na(end) ~ "point",
    str_detect(event_date, "/") ~ "range",
    str_detect(event_date, "\\|") ~ "limit"
  )) |>
  mutate(end = if_else(is.na(end), start, end)) |>
  mutate(start_fuzzy = if_else(str_detect(start, "[%~?]"), TRUE, FALSE)) |>
  mutate(end_fuzzy = if_else(str_detect(end, "[%~?]"), TRUE, FALSE)) |>
  mutate(across(c(start, end), ~str_remove_all(., "[%~?]"))) |>
  mutate(across(c(start, end), ~str_sub(., 1, 4), .names="{.col}_year") ) |>
  mutate(across(c(start_year, end_year), parse_number))  |>
  mutate(year_type = case_when(
    start_year==end_year ~ "single",
    .default = date_type
    ))

  # mutate(when = if_else(date_type=="point", date1, NA)) |>
  # mutate(notBefore = if_else(date_type=="limit", date1, NA),
  #        notAfter = if_else(date_type=="limit", date2, NA)) |>
  # mutate(from = if_else(date_type=="range", date1, NA),
  #        to = if_else(date_type=="range", date2, NA))


events_longer <-
events_xlsx |> 
  mutate(year = parse_number(str_sub(event_date))) |>
  select(event_id, event, rowid, keyword, year, event_date, date_type, date_sort, book1:bookrem) |>
  # add short titles. 
  inner_join(events_short_titles_xlsx |> select(event_id, event_short), by="event_id") |>
  pivot_longer(book1:bookrem, names_to = "book", values_to = "evs") |>
  filter(!is.na(evs)) 

events_evs <-
events_longer |>
  unnest_regex(ev, evs, pattern=" +", drop = F, to_lower = F) |>
  # ev doesn't need the bookX: prefix as that's also in the book col.
  mutate(ev = str_remove_all(ev, "book(\\d|rem):")) |>
  relocate(book, ev) |>
  mutate(xmlid = case_match(
    book,
    "bookrem" ~ "atb-book-of-remembrances",
    "book1" ~ "atb-book-one",
    "book2" ~  "atb-book-two",
    "book3" ~  "atb-book-three"
  )) |>
  inner_join(atb_evs_places_xml_txt |> select(-div_id), by=c("xmlid", "ev"))



events_metadata <-
events_xlsx |> 
  mutate(year = parse_number(str_sub(event_date))) |>
  select(event_id, event, rowid, keyword, year, event_date, date_type, date_sort) |>
  # add short titles. 
  inner_join(events_short_titles_xlsx |> select(event_id, event_short), by="event_id") |>
  # add dates
  inner_join(events_dates |> select(event_id, start_year, end_year, year_type), by="event_id")


data_reactable <-
  events_metadata |>
  select(event_id, event, keyword, start_year, end_year, year_type, date_sort) |>
  arrange(date_sort)
```




```{r}
# see places qmd for details. love it when i can just reuse code and do stuff in an hour!
# needs some renaming for this context though, to diff mentions from places...

atb_place_tag_bkrem_read <- read_html(paste0(get_shared_data(), "xml/places/atb_bookrem_places_table.xml"))
atb_place_tag_bkrem_table <- html_table(atb_place_tag_bkrem_read, na.strings = "")[[1]] 

atb_place_tag_book1_read <- read_html(paste0(get_shared_data(), "xml/places/atb_book1_places_table.xml"))
atb_place_tag_book1_table <- html_table(atb_place_tag_book1_read, na.strings = "")[[1]]

atb_place_tag_book2_read <- read_html(paste0(get_shared_data(), "xml/places/atb_book2_places_table.xml"))
atb_place_tag_book2_table <- html_table(atb_place_tag_book2_read, na.strings = "")[[1]]

atb_place_tag_book3_read <- read_html(paste0(get_shared_data(), "xml/places/atb_book3_places_table.xml"))
atb_place_tag_book3_table <- html_table(atb_place_tag_book3_read, na.strings = "")[[1]]


# reusing quotes code, seems to work.
# do this before filtering anything so you can reuse in different ways when needed.
place_tag_table_all <-
  bind_rows(
    atb_place_tag_bkrem_table |> mutate(src="Book Rem", xmlid="atb-book-of-remembrances"),
    atb_place_tag_book1_table |> mutate(src="Book 1", xmlid="atb-book-one"),
    atb_place_tag_book2_table |> mutate(src="Book 2", xmlid="atb-book-two"),
    atb_place_tag_book3_table |> mutate(src="Book 3", xmlid="atb-book-three")
  ) |>
  # add type for titlePage, then convert titlePage to div
  mutate(type = if_else(element=="titlePage", "titlePage", type)) |>
  mutate(element = if_else(element=="titlePage", "div", element)) |>
  # now you need to fill down the pb id etc, then filter out unwanted rows.
  # with this version you have to make the pb (etc) column first.
  mutate(level = if_else(element %in% c("front", "body", "back"), element, NA)) |>
  mutate(pb_id = if_else(element=="pb", id, NA)) |>
  mutate(div_id = if_else(element%in% c("div", "titlePage"), id, NA)) |>
  mutate(ev_id = if_else(type=="event", id, NA)) |>
  # group by src *before* fill prevents any overspill between Books.
  group_by(src) |>
  # nb empty div tag at the beginning of B2 (before the first pb_id)
  fill(level, pb_id, div_id, ev_id) |>
  # can't use parse_number on div_id, so this makes new per-id num IDs
  # also needs group_by src to restart for each Book, not for the ranking itself
  mutate(divn_id = dense_rank(div_id)) |>
  ungroup() |>
  # numerical page number from id.
  mutate(page = parse_number(str_extract(pb_id, "\\d+"))) |>
  # div can have more than one page; get just the start page (eg for location in reactable)
  # should work for end of div as well? may need to check.
  # needs to group by src as well as div id
  group_by(src, div_id) |>
  mutate(div_start = min(page), div_end=max(page)) |>
  ungroup() |>
  # correction for ev ids  
  # NA -end ids apart from anchor; remove -start or -end suffix from the rest.
  mutate(ev_id = case_when(
    str_detect(ev_id, "-end") & element !="anchor" ~ NA,
    str_detect(ev_id, "-end") ~ str_remove(ev_id, "-end"),
    .default = str_remove(ev_id, "-start")
  )) |>
  # add an ev start page as per div ids.
  # need case_when or if_else because (unlike div) not all rows are in events.
  group_by(src, ev_id) |>
  mutate(ev_start = case_when(
    !is.na(ev_id) ~ min(page)
  )) |>
  ungroup() |>
  # ensure every row has a uid
  rowid_to_column()


#  filter(!place_id %in% c("Engl_094", "Irel_096", "Fran_097"))  
# not filtering out countries... but you've done that in atb_places

# no multis so don't need unnest. 
# first version without grouping
# place_tag_table <-



# you need page numbers rather than ev ids but can link or highlight within page
# book_id isn't the same as the xmlid, need underscores instead of hyphens. 
# https://thornton.kdl.kcl.ac.uk/edition/?p0.do=book_two&p0.lo=p.104&p0.vi=modern&hi=aow1_1595_child_accident
# or highlight the place name?
# https://thornton.kdl.kcl.ac.uk/edition/?p0.lo=p.12&p0.vi=modern&hi=place:Rich_043


place_tag_table_event <-
place_tag_table_all |>
  filter(element=="name") |> # 1439 
  # drop any place without ref (you can't do anything with them anyway)
  filter(str_detect(ref, "place:")) |> # 1402
  #unnest_regex(pers_id, ref, pattern = " +", to_lower = F, drop = F) |> # don't need this
  #relocate(pers_id, .after = rowid) |>
  mutate(ref=str_remove(ref, "place:")) |>
  rename(place_id=ref) |>
  # drop rowid and make this distinct to avoid dups
  distinct(place_id, text, xmlid, src, ev_id, page, div_id) |>
  #select(rowid, place_id, text, xmlid, src, ev_id, page, div_id) |>
  # url for a page needs another format!
  mutate(book_id = case_when(
        src=="Book 1" ~ "book_one",
        src=="Book 2" ~ "book_two",
        src=="Book 3" ~ "book_three",
        src=="Book Rem" ~ "book_rem"
      )) |>
  #rowid_to_column("uid") # don't need to do this though occasionally something needs a 1- sequence. 
  # try another distinct here lol
  inner_join(events_evs |> distinct(ev, xmlid, event_id, event, ev_text_html), by=c("ev_id"="ev", "xmlid")) |>
  # ok at this point you have one row per ev.
  # you want to end up with one row per event+place 
  # BUT retaining per ev formatting and page links
  # so first you need book + page to make urls
  mutate(page_url = glue("https://thornton.kdl.kcl.ac.uk/edition/?p0.do={book_id}&p0.lo=p.{page}&p0.vi=modern&hi=place:{place_id}")) |>
  mutate(page_link_html = glue("<a href='{page_url}' target='_blank'>{src}: {page}</a>")) |>
  # page link makes a heading; it would need to be h5 i think, not sure what that'll look like
  mutate(page_ev_html = glue("<div class='page_ev_html'><h5>{page_link_html}</h5> {ev_text_html}</div>")) |> # 1028
  # to dedup >1 ev-place-page
  distinct(place_id, event_id, page_ev_html) |> # 988
  group_by(place_id, event_id) |>
  summarise(place_event_html = glue_collapse(page_ev_html, sep=" \n"), .groups = "drop_last") |>
  ungroup()




atb_places_geo_event <-
  atb_places |>
  mutate(place_type = case_when(
    level %in% c("geo", "non-country") ~ "geographical",
    level=="subregion" ~ "region",
    .default = level
  ))  |> 
  mutate(country = case_when(
    place_id=="Iris_130" ~ NA,
    is.na(country) ~ "England",
    .default = country
  )) |>
  # prob wont need place segments tbh. or wikidata stuff.
  select(place_id, address= label, place_type, level, type, atb_places_link_html, lat, long) |>
  inner_join(place_tag_table_event, by="place_id") |>
  inner_join(events_metadata, by="event_id") |>
  # should this include fuzziness? it's just so complicated.
  mutate(display_year = case_when(
    start_year==end_year ~ as.character(start_year),
    .default = glue("{start_year}-{end_year}")
  )) |>
  mutate(event_link_html = glue("Search edition for mentions of <a href='https://thornton.kdl.kcl.ac.uk/entities/?hi={event_id}&v=expanded' target='_blank'>this event</a>"))
```



```{r}
ojs_define(ojs_atb_geo_event = atb_places_geo_event)
ojs_define(ojs_events = events_metadata)
```


```{ojs}
ojsAtbGeoEvent = transpose(ojs_atb_geo_event)
ojsEvents = transpose(ojs_events)
```



```{ojs} 
// for the slider
// fortunately, earliest and latest dates are both single points...
earliest = d3.min(ojsAtbGeoEvent.map(d => d.start_year));
latest = d3.max(ojsAtbGeoEvent.map(d => d.end_year));
```



```{ojs} 
function createGeoDataEvent(data) {
  const geoData = {
    type: 'FeatureCollection',
    crs: {type: 'name', properties: {name: 'urn:ogc:def:crs:EPSG::27700'}},
    features: []
  };
  data.map(h => {
    geoData.features.push({
      type: 'Feature',
      properties: {
        // this lacks a uid i think. shouldn't matter
        id: h.place_id,
        event_id: h.event_id,
        address: h.address,
        level: h.level,
        type: h.place_type,
        //category: h.category,
        keyword: h.keyword,
        event: h.event,
        event_year: h.display_year,
        start_year: h.start_year,
        end_year: h.end_year,
        event_link_html: h.event_link_html,
        places_link_html: h.atb_places_link_html,
        event_text_html: h.place_event_html
      },
      geometry: {
        type: 'Point',
        coordinates: [h.long, h.lat]
      }
    });
  });
  return geoData;
}
```




```{ojs}
// array of event descriptions for datalist. hmm maybe. would need to sort out ordering.
// events_list = ojsAtbGeoEvent.map(d => d.event)
// make array of event/place ids for filter
// you should add keywords to the search...

searched_event = search_events.map(d => d.event_id)
searched_place = search_events.map(d => d.place_id) // i think you just want the event id for the search filter


labelEventKeywordData = [
  ...new Set(
    ojsAtbGeoEvent  // data source
      .flatMap((n) => n.keyword )
      .sort()
  )
]
```


## The Map

Features:

* search event descriptions or keywords (table below for reference)
* click on cluster numbers (or zoom in/out) to drill down into areas
* click on single markers to view event details and texts below the map, with links to the DSE.

(*Caveats:* There may be errors, and not all places are of equal significance to Thornton's narratives.)


```{ojs}
viewof search_events = Inputs.search(ojsAtbGeoEvent, {label: "Search events", columns: ["event", "keyword"], width: 400})

/*
viewof rangeYears = Inputs.form([
  Inputs.range([earliest, latest], {label: "From", value: earliest, step: 1}), // start_year and end_year
  Inputs.range([earliest, latest], {label: "To", value: latest, step: 1})]
                         );
*/

//viewof eventKeywordCheck = Inputs.checkbox(labelEventKeywordData, {label: "Event keyword", value: labelEventKeywordData})
```




```{ojs}

map = {
  // Make a container in the Observable DOM to hold the map
  let parent = DOM.element('div', { style: `width:${width}px; height:500px` });
  // Make sure Observable renders the parent container so Leaflet can figure out its size
  yield parent;
  
  let map = L.map(parent)
      .setView([54.3, -3.5], 6); // uk+ire // higher numbers are more zoomed in
  
  // OSM fallback if maptiler fails (or simply slow to load!)
  // possible to customise osm to simply be a bit more subtle?
  // 2026-03-23 weird access blocked messages, though map shows. try reinstating maptiler- that's ok in mindseye.
  
  let osmTileLayer = L.tileLayer( 'http://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png', {
      attribution: '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a>',
      maxZoom: 16,
      minZoom: 5
  })
  .addTo(map);
  
  /*
*/
  const key = 'Qm3IHfglBpG81gp4ZK4G';
  let mtTileLayer = L.tileLayer(`https://api.maptiler.com/maps/base-v4/{z}/{x}/{y}.png?key=${key}`,{ //style URL
    tileSize: 512,
    zoomOffset: -1,
    minZoom: 4,
    attribution: "\u003ca href=\"https://www.maptiler.com/copyright/\" target=\"_blank\"\u003e\u0026copy; MapTiler\u003c/a\u003e \u003ca href=\"https://www.openstreetmap.org/copyright\" target=\"_blank\"\u003e\u0026copy; OpenStreetMap contributors\u003c/a\u003e",
    crossOrigin: true
}).addTo(map);
  

  
  const geojsonData = createGeoDataEvent(ojsAtbGeoEvent);
  
  // tweak clustering...
  
  /*
  Customising the Clustered Markers

how do i just make the text of the markers a bit more prominent without changing anything else?


zoomToBoundsOnClick
whether to turn this off? 
but you have no way of knowing when you've zoomed in far enough for clicking to open spiderfyed singles!

maxClusterRadius i think this def helps.

  */
  
  const mcg = markerCluster({
          spiderfyDistanceMultiplier: 1.8, 
          zoomToBoundsOnClick: true,
          maxClusterRadius: 40
      }).addTo(map);
	
	// todo did i find code to have different markers for types/categories?
	
	// single markers with a simple popup for testing. is L.geoJson the same as L.geoJSON ? seems to be
  let markers = L.geoJSON(null, {
  
  
    onEachFeature: function (feature, layer) {
    
      const data = feature.properties;
      
      // tooltip shows brief info. event short much too short to be useful!
      
      const tip = `<div class="popup">${data.address} (${data.display_year})</div>`;
      
      layer.bindPopup(tip); // click.
      //layer.bindTooltip(tip, {sticky: true}); // hover. 
      
      // make html for fuller details
      
      // to handle NAs (undefined) need to start with an empty const. don't think there are any in this atb data though
      //let linkedlleps = '';
      //if(data.linked_llep) {linkedlleps = `<div><p><i>Examination(s):</i></p> <ul>${data.linked_llep} </ul></div>`; } // don't need else.
      
      
      const html = `<h3>${data.address}</h3>` + 
                    `<h4>${data.event}</h4>` +
                    `<p>${data.event_year}</p>` +
                    `<ul><li>${data.places_link_html}</li><li>${data.event_link_html}</li></ul>` +
                    `<div>${data.event_text_html}</div>` 
                    //`<p>${data.page_link_html}</p>`
                    ; 
                    
      
      //would use textContent for plain text. innerHTML for formatted HTML
      // this needs bindPopup to work nicely on clustered things.
      
      layer.on('click', () => {
        document.getElementById('atb_info').innerHTML = html; 
        })
        
    },
    
    pointToLayer: function (feature, latlng) {
    
    // it works! but comment out because you don't have category any more
    /*const d = feature.properties;
    
    let customColour = "red";
    if (d.category=="home") customColour = "#CC6000";
    else customColour = "#ff7800";
    */
			return L.circleMarker(latlng, {
				radius: 7, 
				//fillColor: customColour,
				fillColor: "#CC6000", // a bit darker than "#ff7800" 
				color: "#000",
				weight: 1, // width of the circle border.
				opacity: 1,
				fillOpacity: 1
			});
     }
     
  }) ;  
  

  
// Once you have populated your Leaflet GeoJSON Layer Group (typically with geojsonLayer.addData(geoJsonObject), then instead of adding that group to your map, simply add it into your MarkerClusterGroup:
  markers.addData(geojsonData).addTo(mcg);

  // This is to make these variables accessible in other Observable cells.
  // It is not necessary for regular HTML web pages.
  
  mutable mapVariables = [mcg, markers];
}

```


<div class="atb_description">

<div id="atb_info"></div>

</div>



```{ojs}

mapFilters = {
  // This handles the changes 
  // In normal HTML environment, this would be inside the slider event handling function.

  mapVariables[1].eachLayer(l => {
    if(
      // single year in data: l.feature.properties.start_year >= rangeYears[0] && l.feature.properties.end_year <= rangeYears[1] 
      // year range in data
      //l.feature.properties.start_year <= rangeYears[1] && l.feature.properties.end_year >= rangeYears[0] 
      //&&
      //eventKeywordCheck.includes(l.feature.properties.keyword)
      searched_event.includes(l.feature.properties.event_id)
    )  
    {
      l.addTo(mapVariables[0]);
    } else {
      l.removeFrom(mapVariables[0]);
    }
  })

  return 'howdy'; // do i need this return line at all? probably not...
}
```

```{ojs}
mutable mapVariables = null;
```



```{ojs}
markerCluster = L, require('leaflet.markercluster@1.1.0').catch(() => L.markerClusterGroup)
```


```{ojs}
markerClusterCSS = html`<link href='${resolve('leaflet.markercluster@1.1.0/dist/MarkerCluster.Default.css')}' rel='stylesheet' />`
```





```{css}
.atb_description { border: 1px solid #d3d3d3; background-color: #f5f5f5; padding: 10px; margin: 10px 5px;}
.atb_description div#atb_info h3 {margin: 20px 0;}
.atb_description div#atb_info div ul li {list-style-type: none;}
.atb_description div#atb_info ul {padding-left: 0;}
.atb_description div#atb_info ul li {margin-left: 20px;}
.atb_description div#atb_info div ul li h4 {margin-top:0;}

span.hi_place {padding:2px; background-color: white;}
```


## Table of Events 

```{r}

## https://glin.github.io/reactable/articles/custom-rendering.html#headers  
## plus https://glin.github.io/reactable/articles/building-twitter-followers.html?q=replace#finishing-touches for replace function
## js function to replace _ in header names with space. often wouldn't need anything else.
headerMakeSpace <- JS("function(column, state) {
      return column.name.replace('_', ' ')
    }")



## filterInput functions ####

## you can probably compare the internal filter functions to the external ones if you want to make adaptations...

# (internal) data list column filter for a table with the given ID
# based on https://glin.github.io/reactable/articles/custom-filtering.html#data-list-filter
# additions:
# sort values alphabetically (default seems to be order of appearance in data)
# placeholder text (default empty)
# might be possible to extend to change type? or better to have separate functions for numeric etc?

dataListFilterInput <- function(tableId, placeholder=NULL, style = "width: 100%; height: 28px;") {
  function(values, name) {
    dataListId <- sprintf("%s-%s-list", tableId, name)
    tagList(
      tags$input(
        type = "text",
        list = dataListId,
        placeholder = placeholder,
        oninput = sprintf("Reactable.setFilter('%s', '%s', event.target.value || undefined)", tableId, name),
        "aria-label" = sprintf("Filter %s", name),
        style = style 
      ),
      tags$datalist(
        id = dataListId,
        # added sort here
        lapply(sort(unique(values)), function(value) tags$option(value = value))
      )
    )
  }
}



# Filter by case-sensitive exact text match. 
# if you don't want case-sensitive you'll probably need regex.
# see also https://github.com/glin/reactable/issues/318#issuecomment-2107704839
# in some examples the final line doesn't have String() - idk whether it makes a difference    
# return row.values[columnId] === filterValue; 

filterTextExact <- JS("function(rows, columnId, filterValue) {
        return rows.filter(function(row) {
          return String(row.values[columnId]) === filterValue
        })
      }")


# dropdown select with "All" as default.

selectFilterInput <- function(tableId, style = "width: 100%; height: 28px;") {
  function(values, name) {
    tags$select(
      # Set to undefined to clear the filter
      onchange = sprintf("Reactable.setFilter('%s', '%s', event.target.value || undefined)", tableId, name),
      # "All" has an empty value to clear the filter, and is the default option
      tags$option(value = "", "All"),
      # added sort here (but what if you have a value alphabetically < All?)
      lapply(sort(unique(values)), tags$option),
      "aria-label" = sprintf("Filter %s", name),
      style = style
    )
  }
}


```



```{r}

reactable(

  data_reactable, 
  
  columns = list(

    
    # linkify a column (using first_img). make sure you use the right data_reactable name!
    # https://thornton.kdl.kcl.ac.uk/entities/?hi={event_id}&v=expanded
    event_id = colDef(name="",
                      maxWidth = 90,
                      filterable = FALSE,
                      html = TRUE,
                      cell = function(value, index){
              sprintf('<a href="https://thornton.kdl.kcl.ac.uk/entities/?hi=%s&v=expanded" target="_blank">DSE</a>', data_reactable$event_id[index])
            }), 
    
    event = colDef(minWidth = 200),
    
    start_year = colDef(name="from", maxWidth = 100),
    end_year = colDef(name="to", maxWidth = 100),
    
    # select dropdown + exact text search
    keyword = colDef(maxWidth = 150 ,
        filterInput = selectFilterInput("data-table"),
        # exact text search to avoid the 'male' in 'female' problem
       filterMethod = filterTextExact
    ),
    
    # # datalist for a single column. nb the function has placeholder and style options
    # keyword = colDef(
    #   filterInput = dataListFilterInput("data-table", style = "width: 100%; height: 32px;") 
    #   ), # noticed default 28px isn't quite as tall as the std box, odd.
    
    # columns to hide
    year_type = colDef(show = FALSE),
    date_sort = colDef(show = FALSE)

),

  ## dropped a few options that I'm unlikely to use/change. (eg rownames)

  # groups are esp useful for wider tables, but take a fair bit of work.
  #   columnGroups = list(
  #     colGroup()
  # ),

  elementId = "data-table", # change if more than one on the page!

  filterable = TRUE,

  defaultColDef = colDef(
      header = headerMakeSpace, # replaces underscore with space. often all you need to make headers presentable!
      minWidth = 120,
      sortNALast = TRUE,
    # start with this commented out as it can slow down large/texty tables. but it's usually worth using.
    ## can exclude specific column(s) by name in the if() eg '& name !="url_text"'
      filterInput = function(values, name) {
        if (is.character(values) ) {
        dataListFilterInput("data-table")(values, name)
      }
    } # /filterInput
   ), # /defaultcoldef

  #defaultColGroup = NULL,

  pagination = TRUE,
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100),
  paginationType = "numbers", # "numbers" for page number buttons, "jump" for a page jump (more compact up-down arrows, useful for tables with a lot of pages)
  showPagination = NULL, # Defaults to TRUE if the table has more than one page.
  showPageInfo = TRUE,

  
# appearance options might vary a bit 
  highlight = TRUE, # highlight on hover?
  outlined = TRUE, # borders around the table? i quite like the borders actually.
  bordered = TRUE, # borders around the table and every cell?
  borderless = FALSE, #  Remove inner borders from table?
  striped = FALSE, # zebra striping for table rows. how well does this go with borders?
  compact = FALSE, # what diff does this make? I think just a bit less padding etc.

  #groupBy = NULL, # colname or c(colnames). aggregation using aggregate arg in colDef()

  #sortable = TRUE,
  #resizable = FALSE,

  #searchable = FALSE, # won't often want this
  #searchMethod = NULL, # custom search method, JS()
  
  # defaultSortOrder = "asc", # "desc" for descending order
  # defaultSorted = NULL, # colname / c(colnames). for different directions use list() list(Species = "asc", Petal.Length = "desc")

  #minRows = 1,
  #paginateSubRows = FALSE, #  When rows are grouped, paginate sub rows? 
  
  #wrap = TRUE, # text wrapping. If TRUE, long text wrapped to multiple lines. If FALSE, text truncated to fit on one line.
    
  fullWidth = TRUE,
  # class = NULL, #   Additional CSS classes to apply to the table.
  # style = NULL, #  Inline styles to apply to table. named list or character string. if named list, property names should be camelCased.
  # rowClass = NULL, # Additional CSS classes to apply to table rows. character string, a JS() function, or an R function
  # rowStyle = NULL, # inline styles for table rows. character string / JS() / R function. if named list, camelCase stuff
  
  theme = getOption("reactable.theme"), #  Theme options for table, specified by reactableTheme(). Can also be a function that returns a reactableTheme() or NULL.
  
  #meta = NULL, #  Custom metadata to pass to JavaScript render functions or style functions. A named list of values that can also be JS() expressions or functions. 

  defaultExpanded = FALSE, # whether to load with nested rows opened 
  

  # details = NULL, # Additional content to display when expanding a row. An R function that takes the row index and column name as arguments, or a JS() function that takes a row info object as an argument. Can also be a colDef() to customize the details expander column.
  
) 

```
















